**Preprinting Microbiology**

### 2017-01-06 submission/Schloss_PrePrints_mBio_2017.Rmd

Many scientists, including microbiologists, have begun to use preprints and other online venues such as social media, blog posts, and videos as methods to engage their peers and non-scientists. Preprints have emerged as a formal venue for microbiologists to disseminate their research. A preprint is an unpublished manuscript that is made publicly available without going through an official peer-review process. Rather, an author can post their manuscript to a preprint server for others to read, share, and make comments. Preprints were initially adopted among physicists in the 1960s as a method of sharing interesting research amongst colleagues. Eventually, this community evolved into what is now the arXiv (pronounced "archive") preprint server that was hosted at the Los Alamos National Laboratories from 1991 to 1999 and then at Cornell University. Among physicists and mathematicians, posting a preprint to arXiv has become the standard publication strategy followed by submission to a peer-reviewed journal. Within those communities researchers have been recognized with international awards for work that has only been posted to arXiv without formal review. Although arXiv has hosted a number of computational biology papers, the server has not drawn widespread attention from biologists. Among the proponents of arXiv, preprints have aided in the development of research communication by accelerating the release of the science and helping it to achieve a wider audience for critique and reception. As adoption of preprints as a component of the scientific method has broadened among biologists, we sought to explore the specific use and concerns of preprints amongst microbiologists.

@@

***Landscape of preprint servers.*** Two preprint servers have sought to parallel the success of arXiv. In 2013, the bioRxiv (pronounced "bio-archive") was launched through Cold Spring Harbor Laboratory as a service to complement arXiv; PeerJ Preprints was initiated at the same time. Both options have similar features: preprint posting is free; each preprint receives a a digital object identifier (DOI) that facilitates the ability to cite preprints in other scholarly work; if the preprint is ever published, the preprint is linked to the published version; the submission process for both options is relatively simple allowing authors to upload a PDF version of their preprint and supplemental materials; preprints are typically posted in under 24 hours; they have built in venues for authors to discuss their research with people who leave comments on the manuscript; preprints undergo a basic screening process to remove submissions with offensive or non-scientific content; and the sites provide article-level metrics indicating the number of times an abstract has been accessed or the preprint has been downloaded. There are several important differences between the two options. First, PeerJ as a for-profit organization and bioRxiv is a non-profit organization. This can be meaningful to authors since some journals, including the ASM Journals, will only accept submissions that have been posted on preprint servers hosted by non-profit organizations. Second, preprints at PeerJ Preprints are posted under the Creative Commons Attribution License (CC-BY) and bioRxiv preprints can be posted under one of four CC-BY licenses or with no permission for reuse. This can be relevant for authors hoping to submit their work a journal as journals will not consider manuscripts posted as preprints under a CC-BY license (e.g. *Proceedings of the National Academy of Sciences*). A cosmetic, but important difference between the two is the layout and feel of the two websites. Compared to the bioRxiv site, the PeerJ Preprint site is more fluid, gives readers the ability to "follow" a preprint, and provides better access to article keywords and the ability to search preprints. With broader acceptance of preprints by traditional journals, many journals, including all of the ASM journals, have established mechanisms to directly submit manuscripts that are posted as preprints on bioRxiv. It is only possible to transfer a PeerJ Preprint for submission to PeerJ. In many ways, preprint servers have taken on the feel of a journal. As adoption of this approach to disseminating research expands, it is likely that the features of these sites will continue to improve. It is also likely that interfaces from third-parties will improve. For example, although Google Scholar includes preprints hosted at bioRxiv and PeerJ Preprints in their search results, PubMed does not. There is hope that the National Institutes of Health (NIH) will renew their interest in including preprints in PubMed search results.

@@

***The use of preprints from an author's perspective.*** Posting research as a preprint gives an author great control over when their work is made public. Under the traditional peer-review model, an author may need to submit and revise their work multiple times over a long period. In contrast, an author can post the preprint at the start of the process for others to consume and comment on as it works its way through the editorial process. A first example illustrate the utility of preprints for improving access to research and the quality of its reporting. In 2014, the Schloss laboratory posted a preprint to PeerJ Preprints describing a method of sequencing 16S rRNA gene sequences using the Pacific Biosciences sequencing platform. At the same time, we submitted for review at PeerJ. While it was under review, we received feedback from someone that read the preprint as well as from Pacific Biosciences that the impact of the results could be limited because of the sequencing chemistry. Instead of resubmitting the manuscript to address the reviews, we generated new data and submitted an updated preprint a year later with a simultaneous submission to PeerJ that incorporated the original reviews as well as the feedback we received from Pacific Biosciences and the independent scientist. In general, the Schloss laboratory has simultaneously posted our manuscripts as preprints and submitted them to a journal. As a second example, this manuscript was posted to bioRxiv as a preprint on XXXXXXX XX, 2017. We then solicited feedback on the manuscript using social media. Two weeks later, we incorporated the comments and posted a revised preprint and submitted the manuscript to *mBio*. During those two weeks the abstract was read XXXX times and the PDF was accessed XXXX times. This process engaged XXXX commenters on bioRxiv, XXXX people on Twitter, XXXX on Facebook, and XXXX via email. We received useful feedback from XXX people. Compared to the two or three scientists that typically review a manuscript, this experience engaged a much larger and more diverse community than had we foregone the posting of a preprint. Although there are concerns regarding the quality of the science posted to a preprint server, we contend that responsible use of preprints as a part of the scientific process can significantly enhance the science.

## Preprints in microbiology
* Rate of use among PP describing microbiology research over time
* Number of revisions
* Number of downloads
* Number of PP published

## Microbiology anecdotes
* MinIon methods
* PacBio paper
* Tardigrade microbiome
* NYC subway microbiome

## Specific challenges for microbiology
* Dual use
* Protection of human subjects
* Conflicts of interest

## Perspective
* Support priority in literature
* Support citation
* Direct submission
* preprint scouts
* mSphereDirect / F1000Prime
* Licensing
* What's in it for ASM Journals and Readers?

>>>>

## Background

Many scientists, including microbiologists, have begun to use preprints and other online venues such as social media, blog posts, and videos as methods to engage their peers and non-scientists. Preprints have emerged as a formal venue for microbiologists to disseminate their research. A preprint is an unpublished manuscript that is made publicly available without going through an official peer-review process. Rather, an author can post their manuscript to a preprint server for others to read, share, and make comments. Preprints were initially adopted among physicists in the 1960s as a method of sharing interesting research amongst colleagues. Eventually, this community evolved into what is now the *arXiv* (pronounced "archive") preprint server that was hosted at the Los Alamos National Laboratories from 1991 to 1999 and then at Cornell University. Among physicists and mathematicians, posting a preprint to *arXiv* has become the standard publication strategy followed by submission to a peer-reviewed journal. Within those communities researchers have been recognized with international awards for work that has only been posted to *arXiv* without formal review. Although *arXiv* has hosted a number of computational biology papers, the server has not drawn widespread attention from biologists. Among the proponents of arXiv, preprints have aided in the development of research communication by accelerating the release of the science and helping it to achieve a wider audience for critique and reception. As adoption of preprints as a component of the scientific method has broadened among biologists, we sought to explore the specific use and concerns of preprints amongst microbiologists.

@@

***Landscape of preprint servers.*** Two preprint servers have sought to parallel the success of arXiv. In 2013, the *bioRxiv* (pronounced "bio-archive") was launched through Cold Spring Harbor Laboratory as a service to complement arXiv; *PeerJ Preprints* was initiated at the same time. Both options have similar features: preprint posting is free; each preprint receives a a digital object identifier (DOI) that facilitates the ability to cite preprints in other scholarly work; if the preprint is ever published, the preprint is linked to the published version; the submission process for both options is relatively simple allowing authors to upload a PDF version of their preprint and supplemental materials; preprints are typically posted in under 24 hours; they have built in venues for authors to discuss their research with people who leave comments on the manuscript; preprints undergo a basic screening process to remove submissions with offensive or non-scientific content; and the sites provide article-level metrics indicating the number of times an abstract has been accessed or the preprint has been downloaded. There are several important differences between the two options. First, PeerJ as a for-profit organization and *bioRxiv* is a non-profit organization. This can be meaningful to authors since some journals, including the ASM Journals, will only accept submissions that have been posted on preprint servers hosted by non-profit organizations. Second, preprints at *PeerJ Preprints* are posted under the Creative Commons Attribution License (CC-BY) and *bioRxiv* preprints can be posted under one of four CC-BY licenses or with no permission for reuse. This can be relevant for authors hoping to submit their work a journal as journals will not consider manuscripts posted as preprints under a CC-BY license (e.g. *Proceedings of the National Academy of Sciences*). A cosmetic, but important difference between the two is the layout and feel of the two websites. Compared to the *bioRxiv* site, the PeerJ Preprint site is more fluid, gives readers the ability to "follow" a preprint, and provides better access to article keywords and the ability to search preprints. With broader acceptance of preprints by traditional journals, many journals, including all of the ASM journals, have established mechanisms to directly submit manuscripts that are posted as preprints on bioRxiv. It is only possible to transfer a PeerJ Preprint for submission to PeerJ. In many ways, preprint servers have taken on the feel of a journal. As adoption of this approach to disseminating research expands, it is likely that the features of these sites will continue to improve. It is also likely that interfaces from third-parties will improve. For example, although Google Scholar includes preprints hosted at *bioRxiv* and *PeerJ Preprints* in their search results, PubMed does not. There is hope that the National Institutes of Health (NIH) will renew their interest in including preprints in PubMed search results.

@@

***The use of preprints from an author's perspective.*** Posting research as a preprint gives an author great control over when their work is made public. Under the traditional peer-review model, an author may need to submit and revise their work multiple times over a long period. In contrast, an author can post the preprint at the start of the process for others to consume and comment on as it works its way through the editorial process. A first example illustrate the utility of preprints for improving access to research and the quality of its reporting. In 2014, the Schloss laboratory posted a preprint to *PeerJ Preprints* describing a method of sequencing 16S rRNA gene sequences using the Pacific Biosciences sequencing platform. At the same time, we submitted for review at PeerJ. While it was under review, we received feedback from someone that read the preprint as well as from Pacific Biosciences that the impact of the results could be limited because of the sequencing chemistry. Instead of resubmitting the manuscript to address the reviews, we generated new data and submitted an updated preprint a year later with a simultaneous submission to PeerJ that incorporated the original reviews as well as the feedback we received from Pacific Biosciences and the independent scientist. In general, our research group has simultaneously posted our manuscripts as preprints and submitted them to a journal. As a second example, this manuscript was posted to *bioRxiv*  as a preprint on **XXXXXXX XX, 2017**. We then solicited feedback on the manuscript using social media. Two weeks later, we incorporated the comments and posted a revised preprint and submitted the manuscript to *mBio*. During that time, the abstract was read **XXXX** times and the PDF was accessed **XXXX** times. This process engaged **XXXX** commenters on bioRxiv, **XXXX** people on Twitter, **XXXX** on Facebook, and **XXXX** via email. We received useful feedback from **XXX** people. Compared to the two or three scientists that typically review a manuscript, this experience engaged a much larger and more diverse community than had we foregone the posting of a preprint. Although there are concerns regarding the quality of the science posted to a preprint server, we contend that responsible use of preprints as a part of the scientific process can significantly enhance the science.

## Preprints in microbiology

***Preprint metrics in microbiology.*** To analyze the use of preprints, we copied the bioRxiv website on December 31, 2016. We chose to analyze the use of *bioRxiv* preprints because these preprints are amenable for submission to ASM journals and there were `r length(biorxiv)` preprints compared to the 2,650 available at *PeerJ Preprint*. Among the `r length(biorxiv)` preprints on bioRxiv, **XXXX** were assigned by the authors into the Microbiology category. When we used a more permissive classification system to identify microbiology-affiliated preprints, we identified **XXXX** additional manuscripts. These microbiology-affiliated preprints were primarily assigned to the **XXXX (N=XX)**, **XXXX (N=XX)**, or **XXXX (N=XX)** categories. One limitation of the *bioRxiv* interface is the inability to assign manuscripts to multiple categories or to tag the content of the preprint. For example, this manuscript could be assigned to either the Microbiology or Scientific Communication and Education categories.

Questions:
* How many preprints are being submitted each month - microbiology category and microbiology-ish
* What fraction of PPs are now published that were deposited before January 1, 2016?
* How many times is each preprint revised (perhaps looking at those that have been published)
* How do PP altmetrics compare to ASM altmetrics over the same time span?
* Where do microbiology/microbiology-affiliated preprints come from?
* How many microbiology/microbiology-affiliated preprints do people submit?

* Distribution of downloads
* Distribution of abstract hits

* Would love to calculate a JIF of preprints with and without those that have not yet been published.

***Microbiology anecdotes.***
* MinIon methods
* PacBio paper
* Tardigrade microbiome
* NYC subway microbiome

***Specific challenges for microbiology.***
* Dual use
* Protection of human subjects
* Conflicts of interest

## Perspective
* Support priority in literature
* Support citation
* Direct submission
* preprint scouts
* mSphereDirect / F1000Prime
* Licensing
* What's in it for ASM Journals and Readers?
* Kills the splash for a paper when it's actually published - author wins, journal loses
